---
title: " "
subtitle: "Good Practice in Scientific Computing 2"
date: "15 March 2023"
output:
  ioslides_presentation:
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Provenance

## Provenance

```{r, out.width='85%', fig.align='center'}
knitr::include_graphics(
  "http://end-to-end-provenance.github.io/provenance_infographic_1_final_juang.jpg"
)

```


## Provenance

```{r, out.width='85%', fig.align='center'}
knitr::include_graphics(
  "http://end-to-end-provenance.github.io/provenance_infographic_2_final_juang.jpg"
)
```

## Tracking provenance

1. Environment and dependencies (e.g. `sessionInfo()` in R)

1. Movement and transformation of data
   - Input data
   - Data manipulation
   - Output files

1. Workflow management systems (Snakemake, Nextflow)

# Consistency

## Data consistency

- Know your minimum information standards!
  - Genomic Standards Consortium: https://www.gensc.org/pages/standards/extensions.html
  - FAIR sharing: https://fairsharing.org/
- Use controlled terminology for interoperability:
  - Gene Ontology
  - Enzyme Commission numbers
  
## Code consistency

- Use popular encoding formats (e.g. ASCII, UTF-8)
- Follow a consistent style
  - Google's style guides: https://google.github.io/styleguide/

# GitHub

## General advice
- Fetch before making changes
- Commit very often
- Archive repos that work
  - Use branches
  - Clone to a backed-up location
- If experimenting, branch/fork the repo
  - Prevents breaking. Don't be the cat that pushed it off the table...

## General advice
- Think about how objects are organised based on use-cases
  - Workshop: Front page, day 1, day 2, ...
  - Documentation: Introduction, step 1, step 2, troubleshooting, FAQ, ...
  - Computational tools: `README.md`, `scr/`, `bin/`, `docs/`, ...

## Collaboration
- Adopt a workflow
  - (small) GitHub flow: https://docs.github.com/en/get-started/quickstart/github-flow
  - (medium) GitLab flow: https://docs.gitlab.com/ee/topics/gitlab_flow.html
  - (large) Microsoft's Release flow: https://devblogs.microsoft.com/devops/release-flow-how-we-do-branching-on-the-vsts-team/
  - General trunk based workflow: https://trunkbaseddevelopment.com/

## Collaboration
- Make small changes in the form of pull requests... PLEASE!
  - Encourages discussion
  - Easier review
- Follow a standard code of conduct
  - Manage expectations 
  - Google's engineering practices: https://github.com/google/eng-practices
- Access control
  - Specific people approve pull requests
  - Merge by pull requests, never just push to main
  
# Thank you!

